---
title: "p8105_hw6_tm2680"
author: "Tatini Mal-Sarkar"
date: "11/21/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(ggridges)
library(purrr)
library(rvest)
library(dplyr)
library(mgcv)
```

# Problem 1
```{r hom_clean}
hom_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

hom_df = 
  read_csv(hom_url) %>% 
  janitor::clean_names() %>% 
  mutate(resolved = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(victim_age),
         city_state = str_c(city, ", ", state),
         victim_race = as.numeric(victim_race != "White"),
         victim_sex = as.numeric(victim_sex == "Male")) %>% 
  filter(city_state != "Dallas, TX" & city_state != "Phoenix, AZ" & city_state != "Kansas City, MO" & city_state != "Tulsa, AL") %>% 
  select(city_state, resolved, victim_age, victim_race, victim_sex)
```

```{r hom_balt_glm}
hom_balt_df = 
  hom_df %>% 
  filter(city_state == "Baltimore, MD")

fit_log = 
  hom_balt_df %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

fit_log %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         lower_bound = exp(estimate - 1.96*std.error),
         upper_bound = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_bound, upper_bound)
```

The adjusted odds ratio for solving homicides comparing non-white to white victims, keeping all other variables fixed, is 0.4406. We are 95% confident the true value for this OR lies between 0.3129 and 0.6204.

```{r glm_func}
glm_func = function(df) {
  log = glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())
  
  broom::tidy(log) %>% 
    mutate(OR = exp(estimate),
         lower_bound = exp(estimate - 1.96*std.error),
         upper_bound = exp(estimate + 1.96*std.error)) %>% 
    filter(term == "victim_race") %>% 
    select(OR, lower_bound, upper_bound)
}
```

```{r hom_glm}
hom_df_glm = hom_df %>% 
  nest(resolved, victim_age, victim_sex, victim_race) %>% 
  mutate(models = map(data, glm_func)) %>% 
  unnest(models) %>% 
  select(-data)
```

```{r hom_glm_plot}
hom_df_glm %>% 
  mutate(city_state = forcats::fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2) + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Solving homicides for non-white vs. white victims",
         x = "City",
         y = "OR")
```

This plot depicts the odds ratio of solving a homicide for a non-white victim compared to a white victim. In cities like Boston, the odds of solving a murder for a non-white victim is 0.118 times the odds of solving a murder for a white victim. We're 95% confident the true value for this OR would lie between 0.049 and 0.285. Because this range doesn't include the null OR value of 1, resolving a homicide and victim race are significantly associated. In other words, in a city like Boston, white victims are likelier to have their case solved. The reverse might hold true in a city like Tampa, where the estimated OR is 1.159. However, because the 95% confidence interval ranges from 0.587 to 2.288, race and resolving a homicide might not be associated. 

# Problem 2
```{r bw_clean}
bw_df = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         mrace = as.factor(mrace),
         missing = sum(is.na(babysex:wtgain)))
```

To build this model, I first narrowed my focus to examining social factors that may influence child birthweight, specifically socioeconomic status and race. To that end, my predictor variables include `fincome` and `mrace`. I recoded `mrace` to white (0) and non-white (1). I hypothesized that higher family income and a white mother lead to higher baby weight.

First, I tried  linear models including each predictor variable individually. 

```{r bw_model_select}
bw_df_mod = bw_df %>% 
  mutate(mrace = as.numeric(mrace != "White"))
  
# Fit of mother's race alone
fit_mrace = lm(bwt ~ mrace, data = bw_df_mod)
summary(fit_mrace)

# Fit of family income alone
fit_fincome = lm(bwt ~ fincome, data = bw_df_mod)
summary(fit_fincome)
```

Then, I tested whether adding maternal race to family income as predictors made a difference, using ANOVA. 
```{r}
# ANOVA of adding maternal race to family income
fit_null_1 = lm(bwt ~ fincome, data = bw_df_mod)
fit_alt_1 = lm(bwt ~ fincome + mrace, data = bw_df_mod)

anova(fit_null_1, fit_alt_1) %>% 
  broom::tidy()
```

Finally, I tested whether adding family income to maternal race as predictors made a difference, using ANOVA once again.
```{r}
# ANOVA of adding family income to maternal race
fit_null_2 = lm(bwt ~ mrace, data = bw_df_mod)
fit_alt_2 = lm(bwt ~ mrace + fincome, data = bw_df_mod)

anova(fit_null_2, fit_alt_2) %>% 
  broom::tidy()
```

